---
title: "Practical Machine Learning Assignment"
author: "Nitish Varshney"
date: "August 22, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Practical Machine Learning Prediction Assignment Writeup

This is Prediction Assignment Writeup. For problem statement and Data sources, see README.md.

**Goal** To predict the manner in which 6 participants did the exercise. This is the "classe" variable in the training set. One should create a report describing how (s)he built model, how (s)he used cross validation, what (s)he think the expected out of sample error is, and why (s)he made the choices for a classifer.

### Model building

**Loading the Data** from the Dataset
```
trainUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testUrl  <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
```

**Processing the Data**
```
set.seed(1234)
```

**Handling Missing Values** : NA option used "na.omit"

### Cross Validation

### Expected out of Sample Error

**Libraries Used**
```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
